\section{System Requirements}
\label{sec:requirements}

When proposing a generic architecture it is fundamental to define the requirements of any system that intents to implement it. These requirements can be of many different types, from energetic , to functional or scalability, all of which will now be mentioned. Any system capable of complying all the presented minimum requirements can make use of the architecture while being sure that their system will be able to properly function as part of a bigger system and with any other compliant systems. 



\begin{description}
\item [Functional requirements] The whole architecture revolves around having a smartphone as the central point for communication. As such the existent functional requirements are all based either on technological requirements  or compatibility with any smartphone. As such the architecture requires the utilisation of smartphone with network capability in order to be able establish communication between smartphone and location or maps server.

Another functional requirement is the one concerning the communication between the beacon component, which provides the smartphone with information about its surroundings and take on many shapes and forms depending on the chosen technology, and the smartphone. This requirement limits the number of usable technologies around the existent sensors. The analysis on the communication compatibility with smartphones has been conducted on section \ref{sec:critical}.

\item [Energetic Requirements] It isn't easy to define energetic requirements of an architecture as it is a very volutile parameter. One of the concerns when making use of smartphone as the central node of communication is the fact that the whole system is constraint on the energetic availability of the smartphone. As such it is of major importance to make sure that the energetic impact of the system when it is working is the smallest possible.

The energetic impact of the system depends on multiple factors such as: Energetic capacity of the smartphone, whose values have been increasing over the years in order to support the higher energetic cost of applications; Cost of the communication between beacon and smartphone, which depends on the size of the messages and the technology; Cost of the communication between smartphone and servers, which depend on the network access available; Cost of the technology's scanning procedure; Cost of map download; Cost of displaying the maps on the smartphone application and finally the overall cost of the application utilisation.

When analysing the costs of each of the previously mentioned factors, the most variable factors are the costs relative to the technology scanning and communication between beacon and smartphone. These cost are dependent on the technology's specific sensor. The remaining communication cost are relatively stable and are independent on the chosen technology. These costs depend mostly on the size of the transmitted messages, which for the ones between smartphone and location server depends on the number of surrounding beacons and for the ones between smartphone and maps server depends on the size of the maps on the maps server.  

The mentioned energetic costs are specific to the cost of one complete operation. The overall cost of the system is dependant on the frequency of the operation. Other than the energy, the used frequency has an impact on the update rate at which the user is capable of knowing its positions. In order to reduce the energetic impact there are algorithms which allow for suspending the process whenever the mobile user is static, i.e. the last location request remains actual without need for update.

Other that the technology specific costs, the energetic costs are highly dependant on the update frequency... 

%NEEDS ACTUAL VALUE???? SAY MAYBE LESS THAN THAT OF THE GPS

\item [Interoperability requirements] The concept behind the proposal of a generic architecture framework is that besides multiple system, which use different technologies, being capable of functioning under the same architecture and interface, they also can work together seamlessly. 

In order to support interoperability, i.e. an indoor building attempting to implement multiple different systems, the smartphone needs to be able to identify the type of input data that it's sending to the location server, while the location server needs to be ready to obtain the user's location through the type of data received.

For the data processing, any two technologies that are compatible with smartphones can be used on the same system. In this system, the user needs to be able of obtain a position through any of the chosen technologies, be it passively or actively, in case that the data collection requires user interaction, and the collected data needs to be tagged and sent to the location server. The system's location server needs to be able to distinguish the two tags and its required to be capable of computing the mobile user's location through either of data types. Since the response from the location server is a location, it is independent of the data provided, which means that the rest of the system doesn't require any changes in the presence of multiple sources of location.

\item [Environmental requisites] The environmental requisites are solely dependant on the system's technology. When studying the possibility of deploying a large scale system which has to be capable of functioning in many different indoor environments, the beacon technology needs to be able to be deployed in those same places.

Another requirement is that the environment where the system is to be deployed needs to have network access, since it is a critical functional requirement. This caution is present in underground environments or other places where mobile network can't be relied upon, which puts the extra requirement of providing another source of network access, such as Wi-Fi.
\end{description}

























