\pdfbookmark{Acknowledgments}{Acknowledgments} 
\begin{acknowledgments}  
 
 
First of all, I would like to thank Prof. João Silva for his availability, willingness to help and openness in the discussions we had during the realization of this thesis. Under his supervision, I had the possibility of working in a focused yet flexible manner, which gave me the possibility of steadily completing this thesis, while also improving myself personally and professionally.  
 
 
I would like to thank my mother and father, for giving me the opportunity of completing this journey. They were always there for me, for both the good and the bad. By putting my happiness above theirs, they were capable of pushing me out of my comfort zone, allowing to discover myself and providing me with experiences that wouldn't be possible otherwise. 
 
 
To my sister, for being much more than just family. Each year, the appreciation for the relation that we have created grows bigger. For all the festivals, conversations and sharing of memories and interests, thank you. 
 
 
A special person that I would like to thank is Laura Berardo. She was always there when I needed to talk or lacked motivation. It's thanks to her that I managed to improve and keep myself focused on my personal and professional objectives. Without her, this and whatever is to come, wouldn't be possible. 
 
 
To my friends, André Malcata, Francisco Leal, André Pombeiro and João Alves, thank you for all the memories that we created together from these last 10 years. You have all contributed so much for the person that I am today.  
 
 
I would like to thank, João Almeida, Lidia Freitas and once again Francisco Leal for their support along this journey. Even though there was almost always someone who was away, we were capable of staying together.  
 
 
\end{acknowledgments} 
\clearpage 
\thispagestyle{empty} 
\cleardoublepage 
 